---
title: "Logistic Regression Vignette"
author: "Ava White, Bukola Ayodele"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Repository web address: 
##### ChatGPT for function outlines/help: https://chat.openai.com/share/460206ba-6cd8-4c6c-8f70-3df76ed08779

# 'logreg.g4' Package
## Logistic Regression Using Numerical Optimization
The functions in this package can be used to perform a logistic regression using numerical optimization, run bootstrapping and generate confidence intervals, plot a fitted regression curve, and generate and plot a confusion matrix with related metrics. We will now go through how to install and use the package, discuss how each function works, and go through examples of each one. 

First, we have to install our package:
```{r}
devtools::install_github("R-4-Data-Science/Final-Project-Group-4/logreg.g4")
library(logreg.g4)
```

To demonstrate how our functions are used, we will first call a dataset to use as an example.


# Function 1- Logistic Regression with Optimization ('logreg')



# Function 2- Least Squares ('leastsquares')



# Function 3- Bootstrapping & Confidence Intervals ('bootstrap_CI')



# Function 4- Plotting a Logistic Curve ('plot_logistic_curve')



# Function 5- Creating a Confusion Matrix & Similar Metrics ('matrix_metrics')



# Function 6- Plotting Model Metrics ('plot_metrics')






